---
title: "Class5"
author: "Kyle Mahowald"
date: "2024-10-03"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
d = read_csv("english.csv")
d$RTms = exp(d$RTlexdec)
```

# Computing the standard error of the mean

First: what does runif() do?

For you to try: play with "samples_drawn"....how is the standard error of the mean affected?

```{r}
# uniform draws
samples_drawn = 10
a = tibble(i=rep(1:100000, samples_drawn))
a = mutate(a, value = runif(n(), 0, 10))

filter(a, i == 1)

# take the means
a.means = group_by(a, i) %>%
  summarize(mean.unif = mean(value),
            sd.unif = sd(value))

ggplot(a.means, aes(x=mean.unif)) + geom_histogram()
sd(a.means$mean.unif)


mean(a.means$sd.unif)/sqrt(samples_drawn)

```

# Standard error of the mean for RTms


```{r}
sd(d$RTms)/sqrt(nrow(d))
```

# Introducing the regression lm()

```{r}
d$RT = d$RTms
l = (lm(data=d, RT ~ Familiarity))

d$Prediction = predict(l)
d$Residuals = d$RT - d$Prediction
summary(d$Residuals)

hist(d$Residuals)
```


# What is the Std.Error in our regression? 

Intuitively, if we sample over and over, how close would we be to the value?

```{r}
l = lm(data=d, RT ~ Familiarity)
summary(l)
d$Resid = residuals(l)

# formula for standard error on the coefficient
sqrt(sum(d$Resid^2)/((nrow(d) - 2) * sum((d$Familiarity - mean(d$Familiarity))^2)))

# bootstrapping
a = NULL
for (i in 1:1000) {
  d.new = sample_n(d, nrow(d), replace=T)
  l = lm(data=d.new, RT ~ Familiarity)
  fam.coef = coef(l)[2]
  a = append(a, fam.coef)
}

hist(a, breaks=50)
sd(a)
summary(l)

-43.577/sd(a)

a

hist(rnorm(1000, 0, 1))
```

# Now scramble the values and see what the coefficient is

```{r}
```

# Now using a continuous variable as a predictor (WrittenFrequency), see how well you can predict RT

```{r}

```

# What is a p-value? 

The probability, under the "null hypothesis", of observing a value as extreme as what we observed.

```{r}
```

# Exercise: Which is better at predicting RT? Familiarity or WrittenFrequency?

```{r}
ggplot(d, aes(x=WrittenFrequency, y=RT)) + geom_point()

summary(lm(data=d, RT ~ Familiarity))
summary(lm(data=d, RT ~ WrittenFrequency))

l.written = lm(data=d, RT ~ WrittenFrequency)
d$resid = residuals(l.written)

ggplot(d, aes(x=WrittenFrequency, resid)) + geom_point()
```



